ABM simulating short-term mating behaviors

### 2022-05-27 submission/manuscript.Rmd
minor language edits to the manuscript

^2^ xxxx

@@

Gender differences in short-term mating behaviors have been a well-documented phenomenon in human sexuality research. Existing studies usually conflate gender differences in mating preferences with differences in sexual behaviors, which is theoretically dubious. Using an agent-based model, the present study investigated the circumstances under which men and women’s differential preferences for short-term mating would result in gender differences in short-term mating behaviors. The results from the model suggested that heterosexual men and women had the same average number of short-term mating experiences and short-term mates even when men had stronger preferences for short-term mating. Gender differences in mating behaviors only emerged when heterosexual men and women in the mating pool (i.e., those with a non-zero number of short-term mating experiences) were considered. Moreover, gay men had a higher average number of both outcomes than lesbian women as well as than heterosexual men when men had stronger preferences for short-term mating. These results suggest that theoretically speaking, even when gender differences in mating preferences existed, those in short-term mating behaviors only occurred among particular populations, or when men’s preferences for short-term mating are not constrained by those of women. Suggestions for future research in human mating psychology and behaviors were provided.

@@

Gender differences in sexual behaviors, especially short-term mating behaviors, have been a well-documented phenomenon in human sexuality research. For example, research has found that heterosexual men, as compared to heterosexual women, have a higher average number of past short-term sexual partners [@oliver_gender_1993; @petersen_meta-analytic_2010; @rissel_heterosexual_2014], have short-term mating more frequently [@petersen_meta-analytic_2010], and engage in extramarital sex more often [@petersen_meta-analytic_2010]. These gender differences have also been observed between gay men and lesbian women [@bailey_effects_1994; @bryant_relationship_1994; @peplau_national_1997; @peplau_sexuality_2004].

In the existing literature, gender differences in sexual behaviors are not usually distinguished from those in attitudes towards or preferences for short-term mating. Some researchers study gender differences on the two levels simultaneously without a conceptual distinction [e.g., @petersen_meta-analytic_2010], whereas some implicitly equate the two, assuming that behavioral differences are a direct expression of the psychological ones [e.g., @schmitt_are_2001].

@@

The present study took gender differences in mating preferences as a presumption, which was conceptualized by sexual strategy theory [@buss_sexual_1993] and repeatedly found support by empirical studies [e.g., @schmitt_universal_2003; @walter_sex_2020]. Using an agent-based model, we investigated whether men and women's differential preferences for short-term mating resulted in any gender differences in short-term mating behaviors, specifically, in the number of short-term mating experiences and short-term mates, and if they did, under what circumstances.

@@

Gender differences in mating preferences have been studied in the light of sexual strategy theory [@buss_sexual_1993]. It posits that since the minimum investment that men devote to their offspring (contribution of sperm through one sexual act) is lower than that of women (gestation, labor, and lactation), men tend to have short-term mating as a larger part of their mating strategy than women do. This is because women's high minimum investment leads to higher opportunity cost (vs. men) if they have sex with a suboptimal partner and give birth to offspring with a low survival chance. As a result, men may have a greater interest in short-term mating, desire a larger number of short-term mates in a given period, and be less selective with respect to accepting a potential mate [@buss_sexual_1993].

The hypotheses derived from sexual strategy theory have received extensive support from empirical studies. For example, men reported to be currently seeking short-term mates to a larger extent than women did [@buss_sexual_1993; @schmitt_are_2001; @schmitt_universal_2003], and a larger proportion of them were in any way seeking short-term mates (vs. not seeking) [@schmitt_universal_2003]. These have been found among U.S. college students [@buss_sexual_1993; @schmitt_are_2001] as well as in ten world regions in a cross-cultural study [@schmitt_universal_2003]. Similarly, men across the world also reported to desire more short-term sexual partners compared to women for a number of given time periods (e.g., a month, a year) [@buss_sexual_1993; @schmitt_universal_2003; @mcburney_preferred_2005]. The gender differences were significant regardless of whether they were estimated by mean [@buss_sexual_1993; @schmitt_are_2001; @schmitt_universal_2003], median [@schmitt_are_2001; @schmitt_universal_2003], or percentage [@schmitt_universal_2003] statistics.

As for mating standards, men are less selective in accepting someone as a potential short-term mate. For example, studies using U.S. college samples found that the minimum percentile ranks that men found acceptable for a potential short-term sexual partner were lower than those accepted by women, both in terms of overall desirability and individual traits (e.g., social status, attractiveness) [@kenrick_evolution_1990; @regan_minimum_1998]. A study also found that when presented with identical descriptions of potential mates, men on average rated them as more desirable than women did [@wiederman_evolution_1998].

Some evidence shows that gender differences in mating preferences also existed between gay men and lesbian women. A survey study using a community sample from the U.S. found that gay men were more interested in short-term mating than lesbian women were, and that this gender difference was comparable to that existing among heterosexual individuals [@bailey_effects_1994]. An indirect piece of evidence on differential standards for short-term mates comes from a recent study finding that significantly more gay men than lesbian women reported to have accepted a casual sexual offer from a same-gender person [@matsick_gender_2021]. Past studies suggest that the gender difference in the acceptance rate of casual sexual offers may originate from men and women’s differential standards for short-term mates [@conley_women_2011; @hald_receptivity_2010]. Thus, as a postulation, gay men may also have lower standards than lesbian women for short-term mates.

## Constraints on men's preferences for short-term mating
Since heterosexual sex involves both a man and a woman in most cases, men's interest in short-term mating can be constrained by women's preference for long-term relationships [@archer_reality_2019; @symons_evolution_1979]. This is because men’s short-term mating preferences can only translate into behaviors when there are women willing to have sex with them. When a new short-term mating encounter occurs, it counts towards men’s total number of short-term mating encounters as well as towards women’s. Similarly, a new short-term mate counts towards both men and women’s total number of short-term mates. Therefore, the total number of short-term mating experiences and short-term mates must be equal between heterosexual men and women on a population level. However, it is important to note that although heterosexual men’s total number of short-term mating is equal to that of women, this does not necessarily mean that the total number of men and women who have ever had short-term mating must be equal. It is possible that there is a smaller proportion of men (vs. the proportion of women) who contribute to the total number of short-term mating. Following this line of reasoning, we would expect to observe no gender difference in short-term mating behaviors among heterosexual individuals even when there were gender differences in mating preferences.

As a comparison, in the cases of gay men and lesbian women, men’s preferences are not constrained by women’s, but only by those of other men, who, arguably, have more similar preferences. This would allow for a more direct behavioral expression of men’s mating preferences. The notion that gay men have less restricted preferences was borne out in the proportion of individuals who have engaged in extradyadic sex [@peplau_sexuality_2004]. A study showed that the proportions of heterosexual men and women who have engaged in extradyadic sex were 26% and 21%, respectively, while the statistics among gay men and lesbian women were 82% and 28%, respectively [@peplau_sexuality_2004]. Therefore, we would theoretically expect to observe gender differences in short-term mating behaviors between gay men and lesbian women. Moreover, we would also expect to find that gay men, as compared to heterosexual men, engaged in more short-term mating behaviors due to the lessened constraint on their mating preferences.

@@

By using a spatial agent-based model, the present study investigated whether men and women’s different preferences for short-term mating resulted in any gender differences in short-term mating behaviors, and if they did, under what circumstances. Preferences for short-term mating were operationally defined as the interest in short-term mating and standards for short-term mates [@buss_sexual_1993]. Short-term mating behaviors were operationally defined as the number of short-term mating experiences and the number of past short-term mates [@petersen_meta-analytic_2010].

We modeled this process among both heterosexual individuals and gay men and lesbian women to examine whether constraints set by the opposite sex’s preferences would affect short-term mating behaviors. Individuals’ sexual orientation was conceptualized in terms of behaviors only in our model. Heterosexual men had short-term mating or formed a long-term relationship only with women, while gay men only with other men, and vice versa for women.

We raised the following hypotheses in the present study: when there were gender differences in preferences for short-term mating, 1) there would be gender differences in short-term mating behaviors among gay men and lesbian women, with gay men being engaged in more such behaviors, and 2) gay men would engage in more short-term mating behaviors as compared to heterosexual men.

@@

We developed an agent-based model to represent a simplified environment where men and women can move around, search for mates, and form long-term and/or short-term relationships. Space and time are modeled as discrete variables. Space was represented as discrete locations on a two-dimensional 33*33 lattice. Agents’ movement in the space was not meant to simulate physical movement but a state of encountering potential mates. Staying at one location represented being committed to a long-term relationship.

The model measured two outcomes: (1) the number of short-term mating experiences of men and women; (2) the number of short-term mates of men and women. Additionally, we also measured the number of men and women in the mating pool (i.e., those with a non-zero number of short-term mating experiences). The average numbers of short-term mating experiences and short-term mates were calculated by taking the average among the whole population of men/women (*N~m~* = *N~w~* = 150) and among those who had engaged in short-term mating.

@@

The interest in short-term mating was modeled as an agent’s likelihood of deciding to have short-term mating at each time step. The standards for short-term mates were modeled as the minimum desirability of a potential mate with whom an agent was willing to have sex. The values of the two parameters were set based on empirical findings in human mating psychology. When there was a gender difference in short-term mating likelihood, men had a 40% likelihood of deciding to engage in short-term mating, while women had a likelihood of 25% [@buss_sexual_1993; @schmitt_are_2001; @schmitt_universal_2003]. When there was no gender difference in short-term mating likelihood, both women and men had a likelihood of 25%. When there was a gender difference in mating standards, men had a mating standard of 3 (the highest possible value = 10), while women had a standard of 5 [@kenrick_evolution_1990; @regan_minimum_1998]. When there was no gender difference in mating standards, both men and women had a standard of 5.

@@

### Procedures
#### Heterosexual procedures.
At each time step, agents first checked whether they were in a long-term relationship. If they were not in a long-term relationship, they set their heading randomly (if they were within the movement range from the birthplace) or faced the birthplace (if they were out of the movement range from the birthplace) and moved by a random distance. The random distance was less than half of the movement range. If they were in a long-term relationship, they did not move. Then, the agents decided on whether to engage in short-term mating by a chance of the chosen likelihood (either 25% or 40%).

Women checked to see if any men were at the same location. If there were, they randomly chose one of them as their potential short-term mate. If both men and women met each other’s mating standards and both of them decided to engage in short-term mating, they had sex. This would result in increasing both parties’ number of short-term mating experiences by one. They also recorded each other on their lists of past short-term mates if they were not on the lists yet.

Then, women randomly selected one man at the same location as their potential long-term partner. There was a 10% of chance that a pair would form a long-term relationship. After forming a long-term relationship, both men and women changed to coupled status and register each other as the long-term partner.

#### Gay men and lesbian women procedures.
At each time step, men and women moved and decided on short-term mating as the agents did in the heterosexual procedures. Half of the agents (75 men and 75 women) were also given the initiator status. The initiators checked to see whether there were other agents at the same location. The rest of the procedures were identical to the heterosexual procedures, except that the agents only chose those with the same gender as potential short-term mates or long-term partners.

@@

## Statistical analysis
Statistical analyses were conducted using `r R.version.string %>% str_replace(" \\(.*\\)", "")` [@r] and the figures were generated by the ggplot2 package [@ggplot2]. Two-tailed independent samples *t*-tests were used for all statistical comparisons. The data were assumed to be normally distributed within each condition but was not formally tested.

@@

The results from Experiment 1 showed that when there were gender differences in preferences for short-term mating, there was no gender difference in the average number of short-term mating experiences among heterosexual individuals (*M~m~* = `r m_exp_all.mean_011`, *SD~m~* = `r m_exp_all.sd_011`; *M~w~* = `r f_exp_all.mean_011`, *SD~w~* = `r f_exp_all.sd_011`), Cohen's *d* = `r exp_all.d_011`. Nor was there a difference in the average number of short-term mates (*M~m~* = `r m_partner_all.mean_011`, *SD~m~* = `r m_partner_all.sd_011`; *M~w~* = `r f_partner_all.mean_011`, *SD~w~* = `r f_partner_all.sd_011`), Cohen's *d* = `r partner_all.d_011` (Figure 1).

\begin{figure}[h]
  \centering
  \includegraphics[width=100mm]{figures/fig1_men_vs_women.pdf}
  \caption{\textbf{Short-term mating behaviors of men and women after 1,000 time steps in the model when gender differences existed in mating preferences.} Violin plots summarizing the outcome variables separately for heterosexual individuals and gay men and lesbian women. Plot (A) shows the average number of short-term mating experiences, and plot (B) shows the average number of short-term mates. Central points show mean values and whiskers represent standard errors, but the standard errors are small and overlap to form a single bar. Statistics were calculated using the full population of men and women in the model (\textit{N\textsubscript{m}} = \textit{N\textsubscript{w}} = 150).}
\end{figure}

@@

In Experiment 1, however, when the means of the outcome variables were calculated among men and women in the mating pool (the number of individuals in the mating pool: *M~m~* = `r m_inpool.mean_011`, *SD~m~* = `r m_inpool.sd_011`; *M~w~* = `r f_inpool.mean_011`, *SD~w~* = `r f_inpool.sd_011`), there were gender differences in short-term mating behaviors when men and women had differential preferences for short-term mating. The average number of short-term mating experiences was higher among heterosexual men (*M* = `r m_exp_ip.mean_011`, *SD* = `r m_exp_ip.sd_011`) than among heterosexual women (*M* = `r f_exp_ip.mean_011`, *SD* = `r f_exp_ip.sd_011`), *t*(`r exp_ip.parameter_011`) = `r exp_ip.statistic_011`, *p* `r exp_ip.p.value_011`, Cohen's *d* = `r exp_ip.d_011`. Likewise, the average number of short-term mates was higher among heterosexual men (*M* = `r m_partner_ip.mean_011`, *SD* = `r m_partner_ip.sd_011`) than among heterosexual women (*M* = `r f_partner_ip.mean_011`, *SD* = `r f_partner_ip.sd_011`), *t*(`r partner_ip.parameter_011`) = `r partner_ip.statistic_011`, *p* `r partner_ip.p.value_011`, Cohen's *d* = `r partner_ip.d_011`.

@@

The results comparing across the two experiments revealed that gay men engaged in short-term mating behaviors more than heterosexual men did (Figure 2). The average number of short-term mating experiences was higher among gay men (*M* = `r m_exp_all.mean_1_11`, *SD* = `r m_exp_all.sd_1_11`) than among heterosexual men (*M* = `r m_exp_all.mean_0_11`, *SD* = `r m_exp_all.sd_0_11`), *t*(`r exp.all.parameter_hg`) = `r exp.all.statistic_hg`, *p* `r exp.all.p.value_hg`, Cohen's *d* = `r exp_all.d_hg`. Similarly, the average number of short-term mates was higher among gay men (*M* = `r m_partner_all.mean_1_11`, *SD* = `r m_partner_all.sd_1_11`) than among heterosexual men (*M* = `r m_partner_all.mean_0_11`, *SD* = `r m_partner_all.sd_0_11`), *t*(`r partner.all.parameter_hg`) = `r partner.all.statistic_hg`, *p* `r partner.all.p.value_hg`, Cohen's *d* = `r partner_all.d_hg`.

\begin{figure}[h]
  \centering
  \includegraphics[width=0.8\columnwidth]{figures/fig2_hetero_vs_gay_men.pdf}
  \caption{\textbf{Short-term mating behaviors of heterosexual and gay men after 1,000 time steps in the model when gender differences existed in mating preferences.} Violin plots summarizing the outcome variables. Plot (A) shows the average number of short-term mating experiences, and plot (B) shows the average number of short-term mates. Central points show mean values and whiskers represent standard errors, but the standard errors are small and overlap to form a single bar. Statistics were calculated using the full population of heterosexual and gay men in the model (\textit{N\textsubscript{hm}} = \textit{N\textsubscript{gm}} = 150).}
\end{figure}

@@

In an exploratory manner, we also compared across conditions to see which of the two dimensions of mating preferences contributed to gender differences in short-term mating behaviors.

@@

Among heterosexual individuals, gender differences in short-term mating behaviors, as calculated among individuals in the mating pool, emerged when men and women had different standards for short-term mates. Heterosexual men (vs. heterosexual women) in the mating pool had a higher average number of short-term mating experiences and short-term mates as long as they had lower mating standards, even when no gender difference existed in short-term mating likelihood (experiences: *M~m~* = `r m_exp_ip.mean_001`, *SD~m~* = `r m_exp_ip.sd_001`, *M~w~* = `r f_exp_ip.mean_001`, *SD~w~* = `r f_exp_ip.sd_001`, Cohen's *d* = `r exp_ip.d_001`; partners: *M~m~* = `r m_partner_ip.mean_001`, *SD~m~* = `r m_partner_ip.sd_001`; *M~w~* = `r f_partner_ip.mean_001`, *SD~w~* = `r f_partner_ip.sd_001`, Cohen's *d* = `r partner_ip.d_001`). In comparison, virtually no gender differences in short-term mating behaviors existed when men and women had the same mating standards (Table 1).

>>>>

^2^ Faculty of Arts, Psychology and Theology, Åbo Akademi University

@@

Gender differences in short-term mating behaviors are well-documented in human sexuality research. Existing studies usually conflate gender differences in mating preferences with differences in sexual behaviors, which is theoretically problematic. Using an agent-based model, we investigated the circumstances under which heterosexual and homosexual men and women’s differential preferences for short-term mating would result in gender differences in short-term mating behaviors. The model showed that when all individuals in a closed heterosexual population were considered,  men and women had the same average number of short-term mating experiences and short-term mates even when men had stronger preferences for short-term mating. Men (vs. women) had a higher average number of both experiences and mates only when heterosexual men and women in the mating pool (i.e., those with a non-zero number of short-term mating experiences) were considered. Moreover, when men had stronger preferences for short-term mating than women, gay men had a higher average number of both experiences and mates compared to both lesbian women and heterosexual men. These results suggest that theoretically speaking, even when gender differences in mating preferences existed, those in short-term mating behaviors only occurred among particular populations, or when men’s preferences for short-term mating are not constrained by those of women. Suggestions for future research in human mating psychology and behaviors were provided.

@@

Gender differences in sexual behaviors, especially short-term mating behaviors, are well-documented in human sexuality research. For example, research has found that heterosexual men, as compared to heterosexual women, have a higher average number of past short-term sexual partners [@oliver_gender_1993; @petersen_meta-analytic_2010; @rissel_heterosexual_2014], and engage in both short-term mating and extramarital sex more often [@petersen_meta-analytic_2010]. These gender differences have also been observed between gay men and lesbian women [@bailey_effects_1994; @bryant_relationship_1994; @peplau_national_1997; @peplau_sexuality_2004].

In the existing literature, gender differences in actual sexual behaviors are not usually distinguished from those in attitudes towards or preferences for short-term mating. Some researchers study gender differences at the two levels simultaneously without a conceptual distinction [e.g., @petersen_meta-analytic_2010], whereas some implicitly equate the two, assuming that behavioral differences are a direct expression of psychological ones [e.g., @schmitt_are_2001].

@@

As a starting point, we took gender differences in mating preferences as conceptualized by the sexual strategy theory [@buss_sexual_1993] and repeatedly supported by empirical investigations [e.g., @schmitt_universal_2003; @walter_sex_2020]. Using an agent-based model, we investigated whether men and women's differential preferences for short-term mating would result in gender differences in short-term mating behaviors, and if so, then under what conditions they would result in differences in the number of short-term mating experiences and short-term mates.

@@

Gender differences in mating preferences have been studied in the light of sexual strategy theory [@buss_sexual_1993]. It posits that since the minimum obligatory investment that men must devote to their offspring (contribution of sperm through one sexual act) is lower than that of women (gestation, labor, and lactation), men tend to be relatively more interested in short-term mating compared to women. This is because the number of offspring women can produce is limited and cannot be increased by mating with a large number of men, whereas the reverse is the case for men. As a result, men are predicted to not only have a greater interest in short-term mating, but also desire a larger number of short-term mates in a given period and be less selective with respect to accepting a potential mate [@buss_sexual_1993].

The hypotheses derived from sexual strategy theory have received extensive support in empirical studies. For example, men have reported to be currently seeking short-term mates to a larger extent than women [@buss_sexual_1993; @schmitt_are_2001; @schmitt_universal_2003], and a larger proportion of men were in any way seeking short-term mates (vs. not seeking) [@schmitt_universal_2003]. Similar results have been found among both U.S. college students [@buss_sexual_1993; @schmitt_are_2001] as well as in ten world regions in a cross-cultural study [@schmitt_universal_2003]. Similarly, men across the world have also reported to desire more short-term sexual partners compared to women over different future time periods (e.g., a month, a year) [@buss_sexual_1993; @schmitt_universal_2003; @mcburney_preferred_2005]. The gender differences have been found regardless of whether they were estimated using mean [@buss_sexual_1993; @schmitt_are_2001; @schmitt_universal_2003], median [@schmitt_are_2001; @schmitt_universal_2003], or percentage [@schmitt_universal_2003] statistics.

As for mating standards, men are less selective in accepting someone as a potential short-term mate. For example, studies using U.S. college samples have found that the minimum percentile ranks that men found acceptable for a potential short-term sexual partner were lower than those accepted by women, both in terms of overall desirability and individual traits (e.g., social status, attractiveness) [@kenrick_evolution_1990; @regan_minimum_1998]. Another study also found that when presented with identical descriptions of potential mates, men on average rated them as more desirable than women did [@wiederman_evolution_1998].

Some evidence shows that gender differences in mating preferences also existed between gay men and lesbian women. A survey study using a community sample from the U.S. found that gay men were more interested in short-term mating than lesbian women were, and that this gender difference was comparable to that existing among heterosexual individuals [@bailey_effects_1994]. An additional indirect piece of evidence on differential standards for short-term mates comes from a recent study finding that significantly more gay men than lesbian women reported to have accepted a casual sexual offer from a same-gender person [@matsick_gender_2021]. Past studies suggest that the gender difference in the acceptance rate of casual sexual offers may originate from men and women’s differential standards for short-term mates [@conley_women_2011; @hald_receptivity_2010]. Thus, as a postulation, gay men may also have lower standards than lesbian women for short-term mates.

## Constraints on men's preferences for short-term mating
Since heterosexual sex in most cases involves one man and one woman, men's relatively high interest in short-term mating can be constrained by women's preference for long-term relationships [@archer_reality_2019; @symons_evolution_1979]. This is because men’s short-term mating preferences – with the relatively rare exception of sexual violence – can only translate into behaviors when there are women willing to have sex with them. When a new short-term mating encounter occurs, it counts both towards men’s total number of short-term mating encounters as well as towards women’s. Similarly, a new short-term mate counts towards both men and women’s total number of short-term mates. Therefore, the total number of short-term mating experiences and short-term mates must be equal between heterosexual men and women at the population level. However, it is important to note that although heterosexual men’s total number of short-term mating is equal to that of women, this does not necessarily mean that the total number of men and women who have ever had short-term mating must be equal. It is possible that there is a smaller proportion of men (vs. the proportion of women) who contribute to the total number of short-term mating. Following this line of reasoning, we would expect to observe no gender difference in short-term mating behaviors among heterosexual individuals even when there are gender differences in mating preferences.

As a comparison, in the cases of gay men and lesbian women, men’s preferences are not constrained by women’s, but only by those of other men, who, arguably, have a similar high interest in short-term mating. This would allow for a more direct behavioral expression of men’s mating preferences. The notion that gay men have less restricted preferences is borne out in the proportion of individuals who have engaged in extradyadic sex [@peplau_sexuality_2004]. A study showed that the proportions of heterosexual men and women who had engaged in extradyadic sex were 26% and 21%, respectively, while the statistics among gay men and lesbian women were 82% and 28%, respectively [@peplau_sexuality_2004]. Therefore, we would expect to observe gender differences in short-term mating behaviors between gay men and lesbian women. Moreover, we would also expect to find that gay men, compared to heterosexual men, engaged in more short-term mating behaviors due to less constraints on their mating preferences.

@@

By using a spatial agent-based model, the present study investigated whether men and women’s differential preferences for short-term mating would result in gender differences in short-term mating behaviors, and if they did, under what circumstances. Preferences for short-term mating were operationally defined as the likelihood of engaging in short-term mating when encountering a mate as well as the standards for short-term mates [@buss_sexual_1993]. Short-term mating behaviors were operationally defined as the number of short-term mating experiences and the number of past short-term mates [@petersen_meta-analytic_2010].

We modeled this process among both heterosexual individuals and gay men and lesbian women to examine whether constraints set by the opposite sex’s preferences would affect short-term mating behaviors. Individuals’ sexual orientation was conceptualized in terms of behaviors only in our model. Heterosexual men had short-term mating or formed a long-term relationship only with women, while gay men only with other men, and mutatis mutandis for women.

We formulated the following hypotheses in the present study: assuming gender differences in preferences for short-term mating, 1) there would be gender differences in short-term mating behaviors among gay men and lesbian women, with gay men being engaged in more such behaviors, and 2) gay men would engage in more short-term mating behaviors as compared to heterosexual men.

@@

We developed an agent-based model to represent a simplified environment where men and women can move around, search for mates, and form long-term and/or short-term relationships. Space and time were modeled as discrete variables. Space was represented as discrete locations on a two-dimensional 33*33 lattice. Agents’ movement in the space was not meant to simulate physical movement but a state of encountering potential mates. Staying at one location represented being committed to a long-term relationship.

The model measured two outcomes: (1) the number of short-term mating experiences of men and women; (2) the number of short-term mates of men and women. Additionally, we also measured the number of men and women in the mating pool (i.e., those with a non-zero number of short-term mating experiences). The average numbers of short-term mating experiences and short-term mates were calculated by taking the average among the whole population of men and women (*N~men~* = *N~women~* = 150) and among those who had engaged in any short-term mating.

@@

The interest in short-term mating was modeled as an agent’s likelihood of deciding to have short-term mating at each time step. The standards for short-term mates were modeled as the minimum desirability of a potential mate with whom an agent was willing to have sex. The values of the two parameters were set based on empirical findings in human mating psychology. When there was a gender difference in short-term mating likelihood, men had a 40% likelihood of deciding to engage in short-term mating, while women had a likelihood of 25% [@buss_sexual_1993; @schmitt_are_2001; @schmitt_universal_2003]. When there was no gender difference in short-term mating likelihood, both women and men had a likelihood of 25%. When there was a gender difference in mating standards, men had a mating standard of 3 (the highest possible value = 10), while women had a standard of 5 [@kenrick_evolution_1990; @regan_minimum_1998], that is, they would not mate with a person with a lower mate value. When there was no gender difference in mating standards, both men and women had a standard of 5.

@@

### Procedures
#### Procedures for heterosexual agents.
At each time step, agents first checked whether they were in a long-term relationship. If they were not in a long-term relationship, they set their heading randomly (if they were within the movement range from the birthplace) or faced the birthplace (if they were out of the movement range from the birthplace) and moved by a random distance. The random distance was less than half of the movement range. If they were in a long-term relationship, they did not move. Then, the agents decided on whether to engage in short-term mating using the predetermined likelihood (either 25% or 40%).

Women checked to see if any men were at the same location. If there were, they randomly chose one of them as a potential short-term mate. If both men and women met each other’s mating standards and both of them decided to engage in short-term mating, they had sex. This would result in increasing both parties’ number of short-term mating experiences by one. They also recorded each other on their respective lists of past short-term mates if they were not on the lists yet.

Then, women randomly selected one man at the same location as their potential long-term partner. There was a 10% of chance that a pair would form a long-term relationship. After forming a long-term relationship, both men and women changed to coupled status and registered each other as their long-term partner.

#### Procedures for gay men and lesbian women agents.
At each time step, men and women moved and decided on short-term mating as the agents did in the heterosexual case. Half of the agents (75 men and 75 women) were also given an initiator status. The initiators checked to see whether there were other agents at the same location. The rest of the procedures were identical to the heterosexual procedures, except that the agents only chose those with the same gender as potential short-term mates or long-term partners.

@@

## Statistical analyses
Statistical analyses were conducted using `r R.version.string %>% str_replace(" \\(.*\\)", "")` [@r] and the figures were generated by the ggplot2 package [@ggplot2]. Two-tailed independent samples *t*-tests were used for all statistical comparisons. The data were assumed to be normally distributed within each condition.

@@

The results from Experiment 1 showed that when there were gender differences in preferences for short-term mating, there was no gender difference in the average number of short-term mating experiences among heterosexual individuals (*M~men~* = `r m_exp_all.mean_011`, *SD~men~* = `r m_exp_all.sd_011`; *M~women~* = `r f_exp_all.mean_011`, *SD~women~* = `r f_exp_all.sd_011`), Cohen's *d* = `r exp_all.d_011`. Nor was there a difference in the average number of short-term mates (*M~men~* = `r m_partner_all.mean_011`, *SD~men~* = `r m_partner_all.sd_011`; *M~w~* = `r f_partner_all.mean_011`, *SD~women~* = `r f_partner_all.sd_011`), Cohen's *d* = `r partner_all.d_011` (Figure 1).

\begin{figure}[h]
  \centering
  \includegraphics[width=100mm]{figures/fig1_men_vs_women.pdf}
  \caption{\textbf{Short-term mating behaviors of men and women after 1,000 time steps in the model when gender differences existed in mating preferences.} Violin plots summarizing the outcome variables separately for heterosexual individuals and gay men and lesbian women. Plot (A) shows the average number of short-term mating experiences, and plot (B) shows the average number of short-term mates. Central points show mean values and whiskers represent standard errors, but the standard errors are small and overlap to form a single bar. Statistics were calculated using the full population of men and women in the model (\textit{N\textsubscript{men}} = \textit{N\textsubscript{women}} = 150).}
\end{figure}

@@

In Experiment 1, however, when the means of the outcome variables were calculated only among men and women in the mating pool (the number of individuals in the mating pool: *M~men~* = `r m_inpool.mean_011`, *SD~men~* = `r m_inpool.sd_011`; *M~women~* = `r f_inpool.mean_011`, *SD~women~* = `r f_inpool.sd_011`), there were gender differences in short-term mating behaviors when men and women had differential preferences for short-term mating. The average number of short-term mating experiences was higher among heterosexual men (*M* = `r m_exp_ip.mean_011`, *SD* = `r m_exp_ip.sd_011`) than among heterosexual women (*M* = `r f_exp_ip.mean_011`, *SD* = `r f_exp_ip.sd_011`), *t*(`r exp_ip.parameter_011`) = `r exp_ip.statistic_011`, *p* `r exp_ip.p.value_011`, Cohen's *d* = `r exp_ip.d_011`. Likewise, the average number of short-term mates was higher among heterosexual men (*M* = `r m_partner_ip.mean_011`, *SD* = `r m_partner_ip.sd_011`) than among heterosexual women (*M* = `r f_partner_ip.mean_011`, *SD* = `r f_partner_ip.sd_011`), *t*(`r partner_ip.parameter_011`) = `r partner_ip.statistic_011`, *p* `r partner_ip.p.value_011`, Cohen's *d* = `r partner_ip.d_011`.

@@

Gay men engaged in short-term mating behaviors more than heterosexual men did (Figure 2). The average number of short-term mating experiences was higher among gay men (*M* = `r m_exp_all.mean_1_11`, *SD* = `r m_exp_all.sd_1_11`) than among heterosexual men (*M* = `r m_exp_all.mean_0_11`, *SD* = `r m_exp_all.sd_0_11`), *t*(`r exp.all.parameter_hg`) = `r exp.all.statistic_hg`, *p* `r exp.all.p.value_hg`, Cohen's *d* = `r exp_all.d_hg`. Similarly, the average number of short-term mates was higher among gay men (*M* = `r m_partner_all.mean_1_11`, *SD* = `r m_partner_all.sd_1_11`) than among heterosexual men (*M* = `r m_partner_all.mean_0_11`, *SD* = `r m_partner_all.sd_0_11`), *t*(`r partner.all.parameter_hg`) = `r partner.all.statistic_hg`, *p* `r partner.all.p.value_hg`, Cohen's *d* = `r partner_all.d_hg`.

\begin{figure}[h]
  \centering
  \includegraphics[width=0.8\columnwidth]{figures/fig2_hetero_vs_gay_men.pdf}
  \caption{\textbf{Short-term mating behaviors of heterosexual and gay men after 1,000 time steps in the model when gender differences existed in mating preferences.} Violin plots summarizing the outcome variables. Plot (A) shows the average number of short-term mating experiences, and plot (B) shows the average number of short-term mates. Central points show mean values and whiskers represent standard errors, but the standard errors are small and overlap to form a single bar. Statistics were calculated using the full population of heterosexual and gay men in the model (\textit{N\textsubscript{hetero}} = \textit{N\textsubscript{gay}} = 150).}
\end{figure}

@@

We also ran analyses across conditions to see which of the two dimensions of mating preferences contributed to gender differences in short-term mating behaviors.

@@

Among heterosexual individuals, gender differences in short-term mating behaviors, when calculated among individuals in the mating pool, emerged when men and women had different standards for short-term mates. Heterosexual men (vs. heterosexual women) in the mating pool had a higher average number of short-term mating experiences and short-term mates as long as they had lower mating standards, even when no gender difference existed in short-term mating likelihood (experiences: *M~men~* = `r m_exp_ip.mean_001`, *SD~men~* = `r m_exp_ip.sd_001`, *M~women~* = `r f_exp_ip.mean_001`, *SD~women~* = `r f_exp_ip.sd_001`, Cohen's *d* = `r exp_ip.d_001`; partners: *M~men~* = `r m_partner_ip.mean_001`, *SD~men~* = `r m_partner_ip.sd_001`; *M~women~* = `r f_partner_ip.mean_001`, *SD~women~* = `r f_partner_ip.sd_001`, Cohen's *d* = `r partner_ip.d_001`). In comparison, virtually no gender differences in short-term mating behaviors existed when men and women had the same mating standards (Table 1).

```{r table_1, include=TRUE, echo=FALSE}
results %>%
  filter(hls %in% c("0_0_0", "0_0_1", "0_1_0", "0_1_1")) %>%
  select(m_exp_ip.mean, m_exp_ip.sd,
         f_exp_ip.mean, f_exp_ip.sd, exp_ip.d,
         m_partner_ip.mean, m_partner_ip.sd,
         f_partner_ip.mean, f_partner_ip.sd, partner_ip.d) %>%
  mutate(
    exp_ip.d = abs(exp_ip.d),
    partner_ip.d = abs(partner_ip.d)
  ) %>%
  mutate(across(where(is.numeric), inline_hook)) %>%
  mutate(
    condition =
      c("Same likelihood*Same standard",
        "Same likelihood*Different standard",
        "Different likelihood*Same standard",
        "Different likelihood*Different standard")
  ) %>%
  column_to_rownames("condition") %>%
  kable(
    digits = 2, format = "latex", booktabs = TRUE,
    caption = "\\textbf{Short-term mating behaviors of heterosexual men and women in the mating pool after 1,000 time steps in the model}",
    row.names = TRUE, col.names = c(rep(c(rep(c("M", "SD"), 2), "d"), 2))
    ) %>%
  row_spec(0, italic = TRUE, align = "c") %>%
  add_header_above(
    c(" " = 1,
      "Men" = 2, "Women" = 2, " " = 1,
      "Men" = 2, "Women" = 2, " " = 1)
    ) %>%
  add_header_above(
    c(" " = 1,
      "Average short-term mating\nexperiences (in-pool)" = 5,
      "Average short-term\nmates (in-pool)" = 5)
    ) %>%
  column_spec(1, width = "2in") %>%
  footnote(
    general = "\\\\textit{Note.} The statistics were calculated with subsamples of men and women who had had engaged in short-term mating behaviors. See Supplementary Materials for the mean and standard deviation of sample sizes.",
    general_title = "", escape = FALSE,
    threeparttable = TRUE, fixed_small_size = TRUE) %>%
  kable_styling(latex_options = "scale_down")
```
